---
title: "Density Analysis"
author: "pthyde123"
date: "2023-11-29"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```


<br>

## Packages
```{r,  echo=TRUE}
library(RColorBrewer)
library(tidyverse)
library(readxl)
library(kableExtra)
```

## Import diversity data file
```{r,  echo=TRUE}
density <- read_excel("data/2023_oat_pea_data.xlsx", 
                      sheet = "density")
```


## Grain (oats and peas)
```{r,  echo=FALSE}

density %>% 
  select(pea_type,oat_grain_g, pea_grain_g,oatName,peaName,
         oatDens,peaDens,oat_percent,pea_percent) %>% 
  mutate(total_grain = (oat_grain_g + pea_grain_g))%>%
  mutate(inter_crop = str_c(oatName,"_",peaName)) %>%
  mutate(density_oat_pea = str_c("oat_",oat_percent,"_","pea_",pea_percent)) %>%

  filter(density_oat_pea != "oat_NA_pea_NA") %>% 
  mutate(density_oat_pea = fct_relevel(density_oat_pea, "oat_0_pea_50","oat_0_pea_100",
                                       "oat_50_pea_0","oat_100_pea_0",
                          "oat_50_pea_50",
                          "oat_50_pea_100", "oat_100_pea_50","oat_100_pea_100"   
                          )) %>% 
  
  ggplot(aes(density_oat_pea, total_grain))+
  geom_boxplot()+
  xlab("density_oat_pea")+
  ylab("Grain g")+
  scale_fill_brewer(palette="Dark2")+
  theme_classic()+
  theme(axis.text.x=element_text(angle = 90, hjust = 1))+
  theme(axis.text=element_text(size=20), axis.title=element_text(size=20,face="bold"))+
  theme(axis.text.x = element_text(face = "bold", color = "black", size = 16))+
  theme(axis.text.y = element_text(face = "bold", color = "black", size = 16))


```




## Genotype Density Interactions

```{r,  echo=FALSE}
density_df <-
density %>% 
  select(pea_type,oat_grain_g, pea_grain_g,oatName,peaName,
         oatDens,peaDens,oat_percent,pea_percent) %>% 
  mutate(total_grain = (oat_grain_g + pea_grain_g))%>%
  mutate(inter_crop = str_c(oatName,"_",peaName)) %>%
  mutate(density_oat_pea = str_c("oat_",oat_percent,"_","pea_",pea_percent)) %>%
  
  filter(density_oat_pea != "oat_NA_pea_NA") %>% 
  mutate(density_oat_pea = fct_relevel(density_oat_pea, "oat_0_pea_50","oat_0_pea_100",
                                       "oat_50_pea_0","oat_100_pea_0",
                                       "oat_50_pea_50",
                                       "oat_50_pea_100", "oat_100_pea_50","oat_100_pea_100"   
  ))


density_df  %>% 
  kable(align = "c") %>% 
  kable_classic(full_width=F,position = "left")%>% 
 scroll_box(width = "900px", height = "300px")



```



<br/>  
<br/> 

## Oat density on pea density interaction

<br/>  

$$
{yield} = \beta_{0} + \beta_{1}({oatName}) + \beta_{2}({peaName}) + \beta_{3}({pea\_percent}) + \beta_{4}({oat\_percent}) + \beta_{5}({oat\_percent}*{pea\_percent}) + \epsilon
$$






```{r,  echo=FALSE}


lm_density  <- lm((total_grain) ~ oatName + peaName + pea_percent + oat_percent+
                  oat_percent * pea_percent,data=density_df)

anova(lm_density)

lm_density  <- lm((oat_grain_g) ~ oatName + peaName + pea_percent + oat_percent+
                  oat_percent * pea_percent,data=density_df)

anova(lm_density)

lm_density  <- lm((pea_grain_g) ~ oatName + peaName + pea_percent + oat_percent+
                  oat_percent * pea_percent,data=density_df)

anova(lm_density)








```













<br/>  
<br/> 

## Oat genotype by pea density interaction 

<br/>  

$$
{yield} = \beta_{0} + \beta_{1}({oatName}) + \beta_{2}({peaName}) + \beta_{3}({pea\_percent}) + \beta_{4}({oat\_percent}) + \beta_{5}({oatName}*{pea\_percent}) + \epsilon
$$



```{r,  echo=FALSE}


lm_density  <- lm((total_grain) ~ oatName + peaName + pea_percent + oat_percent + oatName * pea_percent ,data=density_df)

anova(lm_density)


lm_density  <- lm((oat_grain_g) ~ oatName + peaName + pea_percent + oat_percent + oatName * pea_percent ,data=density_df)

anova(lm_density)


lm_density  <- lm((pea_grain_g) ~ oatName + peaName + pea_percent + oat_percent + oatName * pea_percent ,data=density_df)

anova(lm_density)




```



<br/>  
<br/> 

## Pea genotype by oat percent interaction

<br/>
$$
{yield} = \beta_{0} + \beta_{1}({oatName}) + \beta_{2}({peaName}) + \beta_{3}({pea\_percent}) + \beta_{4}({oat\_percent}) + \beta_{5}({peaName}*{oat\_percent}) + \epsilon
$$


```{r,  echo=FALSE}


lm_density  <- lm((total_grain) ~ oatName + peaName + pea_percent + oat_percent + peaName * oat_percent ,data=density_df)

anova(lm_density)


lm_density  <- lm((oat_grain_g) ~ oatName + peaName + pea_percent + oat_percent + peaName * oat_percent ,data=density_df)

anova(lm_density)


lm_density  <- lm((pea_grain_g) ~ oatName + peaName + pea_percent + oat_percent + peaName * oat_percent ,data=density_df)

anova(lm_density)




```


## Model for plot that clearly shows pea density X pea type interaction

```{r,  echo=FALSE}

density %>% 
  select(pea_type,oat_grain_g, pea_grain_g,oatName,peaName,
         oatDens,peaDens,oat_percent,pea_percent) %>% 
  mutate(total_grain = (oat_grain_g + pea_grain_g))%>%
  mutate(inter_crop = str_c(oatName,"_",peaName)) %>%
  mutate(density_oat_pea = str_c("oat_",oat_percent,"_","pea_",pea_percent)) %>%
  
  filter(density_oat_pea != "oat_NA_pea_NA") %>% 
  mutate(density_oat_pea = fct_relevel(density_oat_pea, "oat_0_pea_50","oat_0_pea_100",
                                       "oat_50_pea_0","oat_100_pea_0",
                                       "oat_50_pea_50",
                                       "oat_50_pea_100", "oat_100_pea_50","oat_100_pea_100"   
  )) %>%
  filter(density_oat_pea %in% c("oat_50_pea_50",
                                       "oat_50_pea_100", 
                                       "oat_100_pea_50",
                                       "oat_100_pea_100")) %>% 
  
  ggplot(aes(density_oat_pea, total_grain))+
  geom_boxplot()+
  xlab("density_oat_pea")+
  ylab("Total Harvest g")+
  scale_fill_brewer(palette="Dark2")+
  theme_classic()+
  theme(axis.text.x=element_text(angle = 90, hjust = 1))+
  theme(axis.text=element_text(size=20), axis.title=element_text(size=20,face="bold"))+
  theme(axis.text.x = element_text(face = "bold", color = "black", size = 16))+
  theme(axis.text.y = element_text(face = "bold", color = "black", size = 16))+
  facet_wrap(~pea_type)



```







<br/>  
<br/> 

$$
{yield} = \beta_{0} + \beta_{1}({oatName}) + \beta_{2}({pea\_type}) + \beta_{3}({oat\_percent}) + \beta_{4}({pea\_percent}) + \beta_{5}({pea\_type}*{pea\_percent}) + \epsilon
$$




```{r,  echo=FALSE}

density_no_0 <- density_df %>% 
  filter(density_oat_pea %in% c("oat_50_pea_50",
                                       "oat_50_pea_100", 
                                       "oat_100_pea_50",
                                       "oat_100_pea_100")) 

lm_density  <- lm((total_grain) ~ oatName + pea_type + oat_percent + pea_percent + 
                  pea_type * pea_percent,data=density_no_0)

anova(lm_density)

lm_density  <- lm((oat_grain_g) ~ oatName + pea_type + oat_percent + pea_percent + 
                  pea_type * pea_percent,data=density_no_0)

anova(lm_density)

lm_density  <- lm((pea_grain_g) ~ oatName + pea_type + oat_percent + pea_percent + 
                  pea_type * pea_percent,data=density_no_0)

anova(lm_density)



```







## Histograms of yield data

```{r, out.width="200px", out.height="200px",echo=FALSE}

density_df %>% 
  ggplot(aes(total_grain))+
  geom_histogram(binwidth=50, color="black", fill="grey")+
  xlab("Yield g")+
  ylab("Count")+
  scale_fill_brewer(palette="Dark2")+
  theme_classic()+
  theme(axis.text.x=element_text(angle = 90, hjust = 1))+
  theme(axis.text=element_text(size=20), axis.title=element_text(size=20,face="bold"))+
  theme(axis.text.x = element_text(face = "bold", color = "black", size = 16))+
  theme(axis.text.y = element_text(face = "bold", color = "black", size = 16))

density_df %>% 
  ggplot(aes(oat_grain_g))+
  geom_histogram(binwidth=50, color="black", fill="grey")+
  xlab("Oat g")+
  ylab("Count")+
  scale_fill_brewer(palette="Dark2")+
  theme_classic()+
  theme(axis.text.x=element_text(angle = 90, hjust = 1))+
  theme(axis.text=element_text(size=20), axis.title=element_text(size=20,face="bold"))+
  theme(axis.text.x = element_text(face = "bold", color = "black", size = 16))+
  theme(axis.text.y = element_text(face = "bold", color = "black", size = 16))

density_df %>% 
  ggplot(aes(pea_grain_g))+
  geom_histogram(binwidth=50, color="black", fill="grey")+
  xlab("Pea g")+
  ylab("Count")+
  scale_fill_brewer(palette="Dark2")+
  theme_classic()+
  theme(axis.text.x=element_text(angle = 90, hjust = 1))+
  theme(axis.text=element_text(size=20), axis.title=element_text(size=20,face="bold"))+
  theme(axis.text.x = element_text(face = "bold", color = "black", size = 16))+
  theme(axis.text.y = element_text(face = "bold", color = "black", size = 16))

```


Yield data looks somewhat normally distributed





## Biomass

```{r,  echo=FALSE}

density %>% 
  
  select(`pea_type`,peaName,Biomass_1_pea_g,Biomass_1_oat_g,
         Biomass_2_pea_g,Biomass_2_oat_g,
         Biomass_3_pea_g,Biomass_3_oat_g,
         Biomass_4_pea_g,Biomass_4_oat_g) %>% 
  filter(pea_type != "na") %>%
  pivot_longer(Biomass_1_pea_g:Biomass_4_oat_g, names_to = "biomass", 
               values_to = "biomass_g") %>% 
  mutate(biomass = as.factor(biomass)) %>%
  mutate(biomass_g = as.numeric(biomass_g)) %>%
  mutate(biomass = fct_relevel(biomass, "Biomass_1_pea_g",
                               "Biomass_2_pea_g","Biomass_3_pea_g","Biomass_4_pea_g",
                               "Biomass_1_oat_g","Biomass_2_oat_g",
                               "Biomass_3_oat_g",
                               "Biomass_4_oat_g")) %>%
  
  ggplot(aes(biomass,`biomass_g`,fill=biomass))+
  geom_boxplot() +
  xlab("Pea Type")+
  ylab("Biomass g")+
  scale_fill_brewer(palette="Dark2")+
  theme_classic()+
  theme(axis.text.x=element_text(angle = 90, hjust = 1))+
  theme(axis.text=element_text(size=20), axis.title=element_text(size=20,face="bold"))+
  theme(axis.text.x = element_text(face = "bold", color = "black", size = 16))+
  theme(axis.text.y = element_text(face = "bold", color = "black", size = 16))




```






```{r,echo=FALSE}

density %>% 
  
  select(`pea_type`,peaName,Biomass_1_pea_g,Biomass_1_oat_g,
         Biomass_2_pea_g,Biomass_2_oat_g,
         Biomass_3_pea_g,Biomass_3_oat_g,
         Biomass_4_pea_g,Biomass_4_oat_g) %>% 
  filter(pea_type != "na") %>%
  pivot_longer(Biomass_1_pea_g:Biomass_4_oat_g, names_to = "biomass", 
               values_to = "biomass_g") %>% 
  mutate(biomass = as.factor(biomass)) %>%
  mutate(biomass_g = as.numeric(biomass_g)) %>%
  mutate(biomass = fct_relevel(biomass, "Biomass_1_pea_g",
                               "Biomass_2_pea_g","Biomass_3_pea_g","Biomass_4_pea_g",
                               "Biomass_1_oat_g","Biomass_2_oat_g",
                               "Biomass_3_oat_g",
                               "Biomass_4_oat_g")) %>%
  
  ggplot(aes(`pea_type`,`biomass_g`,fill=biomass))+
  geom_boxplot() +
  xlab("Pea Type")+
  ylab("Biomass g")+
  scale_fill_brewer(palette="Dark2")+
  theme_classic()+
  theme(axis.text.x=element_text(angle = 0, hjust = 1))+
  theme(axis.text=element_text(size=20), axis.title=element_text(size=20,face="bold"))+
  theme(axis.text.x = element_text(face = "bold", color = "black", size = 16))+
  theme(axis.text.y = element_text(face = "bold", color = "black", size = 16))






```





